Merging Datasets Through Deep learning

### 2018-10-21 old/paper2018/introduction.tex
reorgonized directory

Merging datasets is a key operation for data analytics. A frequent requirement for merging is a join between two columns that frequently have different surface representations for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}, as shown in Table~\ref{table-example}.  This problem occurs for many entity types such as people's names, company names, addresses, product descriptions, conference venues, or even people's faces.  Data management systems have however, largely focussed solely on equi-joins, where string or numeric equality determines which rows should be joined, because such joins are efficient.

Extensions in data management systems for handling `fuzzy' joins across different surface forms for the same entity typically use string similarity algorithms such as edit-distance, Jaro-Winkler and TF-IDF (e.g., \cite{Cohen2003}).  In applying fuzzy joins, a key problem is to avoid quadratic comparisons between pairs of strings drawn from the two columns.  A blocking strategy is used to reduce the number of pairs for fuzzy joins, such that only strings that share a common prefix, or suffix will be compared further.  String matching algorithms often do not work however, because transformations of entity names in perfectly valid ways can yield very different strings.  As shown in Table~\ref{table-example}, \textit{John Smith} is more similar to \textit{John Adams} than \textit{Adams, John}, but of course it is \textit{Adams, John} that is the valid alternate surface form for \textit{John Adams}.

@@

Specifically, our solution to the join problem involves building a system that performs deep metric learning.  That is, the network learns is to map input vectors corresponding to same entity closer in vector space compared to vectors that correspond to different entities.  For the closely related problem of face recognition and person re-identification, deep learning networks have been used successfully to discriminate pairs that represent the same entity from pairs that do not.  However, simply building a model that discriminates pairs of entities is not sufficient; if used naively, this would mean a quadratic comparison between cell values in the two columns to be joined.  We know of no work that has adapted these techniques to the problem of how to use such deep learning models to perform joins efficiently for different surface forms of the same entity.

As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  Entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the problem of matching entity names with a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Douglas Noel Adams}) and a negative element (e.g., John Adams), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.

>>>>

Merging datasets is a key operation for data analytics.  A frequent
requirement for merging is joining across columns that have
different surface forms for the same entity (e.g., the name of a
person might be represented as \textit{Douglas Adams},
\textit{Adams, Douglas} or \textit{Douglas Noel Adams}).  Similarly,
ontology alignment can require recognizing distinct surface forms of
the same entity, especially when ontologies are independently
developed.  This problem occurs for many entity types such as people's names, company names, addresses, product descriptions, conference venues, or even people's faces.  Data management systems have however, largely focussed solely on equi-joins, where string or numeric equality determines which rows should be joined, because such joins are efficient.

Extensions in data management systems for handling `fuzzy' joins across different surface forms for the same entity typically use string similarity algorithms such as edit-distance, Jaro-Winkler and TF-IDF (e.g., \cite{Cohen2003}).  In applying fuzzy joins, a key problem is to avoid quadratic comparisons between pairs of strings drawn from the two columns.  A blocking strategy is used to reduce the number of pairs for fuzzy joins, e.g. only strings that share a common prefix or suffix will be compared further.  But such strategies based on strings often do not work because transformations of entity names in perfectly valid ways can yield very different strings.  As shown in Table~\ref{table-example}, \textit{John Smith} is more similar to \textit{John Adams} than \textit{Adams, John}, but of course it is \textit{Adams, John} that is the valid alternate surface form for \textit{John Adams}.

@@

Specifically, our solution to the join problem involves building a system that performs deep metric learning: the network learns to map input vectors corresponding to same entity closer in vector space than vectors corresponding to different entities.  For the closely related problem of face recognition and person re-identification, deep learning networks have been used successfully to discriminate pairs that represent the same entity from pairs that do not.  However, simply building a model that discriminates pairs of entities is not sufficient; if used naively, this would mean a quadratic comparison between cell values in the two columns to be joined.  We know of no prior work that has adapted these techniques to the problem of how to use such deep learning models to perform joins efficiently for different surface forms of the same entity.

As in face recognition, our system for metric learning is built by training a `triplet siamese network' to learn to map different surface forms of the same entity closer in vector space than different forms of the same entity.  Entity names have very different properties from faces, so we needed to adapt the approaches used in face recognition to the problem of matching entity names.  We use a novel method to choose the sample for training, and novel loss functions as objectives for the metric learning problem.  At training, the network is given a triplet consisting of an `anchor' (e.g. \textit{Douglas Adams}), a positive element (e.g. \textit{Adams, Douglas}) and a negative element (e.g., \textit{John Adams}), and it learns to produce a small distance estimate for the positive pair (close to 0), and a large distance estimate for the negative pair, using an objective function. A classification form of this function that simply learns whether two names belong to the same entity can be used to perform fuzzy joins, but then we would need a blocking strategy analogous to what has been used in the literature to identify the set of joinable pairs \cite{auto-join-joining-tables-leveraging-transformations}.
